General constraints on magic state protocols from the statistical mechanics of \\Wigner negativity

### 2021-06-28 magic_paper.tex
red comments to be addressed

\begin{figure}[t]
    \centering
    \includegraphics[scale=0.35]{figs/distill_bounds.pdf}
    \caption{\textbf{Magic distillation bounds from majorization of Wigner quasi-distributions.} The above plot shows distillation upper bounds for qutrits obtained from majorisation on unital protocols. The constraints are plotted as a function of the depolarizing noise level $\epsilon$ for a noisy Strange state. The bound $R_\infty$ follows from a Lorenz curve analysis, while the bound $R_{10}$ comes from single-shot entropies on the Wigner quasi-distributions, as discussed in Section~\ref{section:entropies}. Numerical bounds $R_{\rm{num}}$ are obtained by using the full set of majorization constraints. The new bounds $R_\infty$, $R_{10}$, $R_{\rm{num}}$ all provide stricter constraints than mana and max--thauma $\theta_{\rm{max}}$ bounds for noisy Strange states.
    }
    \label{fig:distill_bounds}
\end{figure}

@@

\subsection{Magic theories for quantum computation}
\label{sec:mono}
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3]{figs/protocol_diagram.pdf}
    \caption{\textbf{Probing magic protocols with reference states.} 
	A magic protocol (top figure) converts $n$ copies of some noisy magic state $\rho$ with noise parameter $\epsilon$ to $m$ copies of some less noisy state $\rho'$ with $\epsilon' < \epsilon$. The physics of the protocol can be analysed by considering how it would transform some distinguished reference stabilizer state $\tau^{\otimes n}$ (bottom figure). For the case where this reference state is a thermal state at some temperature $T$ our magic distillation bounds are functions of $T$, the noise parameters $\epsilon, \epsilon'$, and the free energy the protocol channel $\E$ injects in. In addition, a quantity $\phi$ (Equation~\ref{eq:phi}) links the computational and physical aspects, and corresponds to the degree to which Wigner negativity in $\rho$ can be assigned a sharp energy via the system Hamiltonian.}
    \label{fig:sketch}
\end{figure}
A number of magic theories exist, where magic states are computational resource states with respect to a natural class of quantum operations that are considered free~\cite{Gour_2019}. One very natural class of free operations are those obtained from Clifford unitaries, Pauli measurements and the ability to discard quantum systems. However, there are also other candidates~\cite{cit:ahmadi, cit:seddon, Wang_2019}.

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.35]{figs/lc_strange.pdf}
    \caption{\textbf{A `heretical' family of Lorenz curves.} Traditionally, Lorenz curves for distributions are monotone increasing cumulant functions that reach a maximum value of $1$. In contrast, Lorenz curves for magic states break through the value of $L(x)=1$ due to the presence of negativity in the associated quasi-probability distributions. The above family of curves correspond to multiple copies of noisy Strange states $\rho\coloneqq\rho_{\rm{S}}(\epsilon)^{\otimes n}$ (\cref{eq:noisysn}) for $n=2,4,6$ within $\R_{\id/3}$. Solid lines represent pure Strange states, while dashed lines represent $\epsilon$-noisy Strange states with depolarising error $\epsilon = 0.1$.
    }
    \label{fig:lcs}
\end{figure}

@@

\begin{figure}[t!]
    \centering
    \subfigure[][]{%
    \label{fig:noCliff}%
    \includegraphics[scale=0.35]{figs/rate_scatter.pdf}}\\
    \subfigure[][]{%
    \label{fig:Cliff}%
    \includegraphics[scale=0.35]{figs/rate_scatter_cliff.pdf}}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.} Shown are two contour plots of the bound on $R(\epsilon, \beta)$ for the case $H = H' = \sum_{k \in \mathbb{Z}_3} k \ketbra{k}$ and $\epsilon' = 0$, where $\beta$ is the inverse temperature and $\epsilon$ is the depolarising noise of the input magic state. 
The top figure \subref{fig:noCliff} does not use any changes of Clifford basis, and the form of the bound depends on both the noise parameter and temperature. The curved dashed line is $\epsilon_\star(\beta)$ is given by Equation \ref{eq:threshold} and $\beta_\star = (kT_\star)^{-1}$ is given by $E_2-\phi = kT_\star \ln 2$. 
In the bottom figure \subref{fig:Cliff} Clifford processing is used resulting in a smoother bound. In both figures the $\beta = 0$ line correspond to the unital bounds.}
    \label{fig:rate_contour}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.35]{figs/R_vs_A.pdf}
    \caption{\textbf{Variation of distillation bounds with Hamiltonian.} We consider $(\epsilon, \epsilon', \beta) = (0.1, 0.0, 0.2)$, and look at a family of Hamiltonians for the qutrit systems, given by $H = A_{\bmo}$ and $H' = (1-p-q)A_{\bmo}+pA_{(1,2)}+q{\rm{diag}}(0,1,2)$.
    }
    \label{fig:rvsa}
\end{figure}

@@

We can now apply this to obtain a family of magic distillation bounds based on R\'{e}nyi divergences.
\begin{theorem}\label{thm:entropic_bounds}
	Consider a general magic state distillation process on prime dimension qudits, that converts a magic state $\rho^{\otimes n} \longrightarrow \E(\rho^{\otimes n})=\rho'^{\otimes m}$and let $\tau$ be any full-rank stabilizer reference state on a qudit. Then the distillation rate $R := m/n$ is upper bounded as,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)}{D^{\rm reg}_\alpha( \rho', \tau')},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and with the regularized divergence
	\begin{equation}
D^{\rm reg}_\alpha( \rho', \tau') := \frac{1}{m} D_\alpha (W_{\rho'^{\otimes m}} || W_{\tau'_m}),
\end{equation}
between the output magic state $\rho'^{\otimes m}$ and $\tau'_m = \E(\tau^{\otimes n})$.
\end{theorem}
The proof is given in Appendix~\ref{app:single-shot}.

@@

If correlations in $\tau'_m$ between subsystems can be neglected then we simply get
\begin{equation}
D^{\rm reg}_\alpha( \rho', \tau') = D_\alpha (W_{\rho'} || W_{\tau'}),
\end{equation}
which gives a generalized version of Theorem \ref{thm:free-energy}.

@@

The Wigner distribution of state $\tau$ is then given by
\begin{align}
	\W[\bmz]{\tau} &= \sum\limits_{k=0}^2 \frac{e^{-\beta E_k}}{\Z}W_{|k\>\<k|}(x, p) \nonumber\\
	&= \sum\limits_{k=0}^2 \frac{e^{-\beta E_k}}{\Z} \delta_{x, k} = \frac{e^{-\beta E_x}}{3\Z},
\end{align}
where $x$ labels one of the three vertical lines in the phase space.
The rescaled Wigner distribution $W_{\rho|\tau}(\bmz)$ can now be computed. It has $9$ components, but these come with multiplicities. In total, there are four distinct values on the phase space, as illustrated in~\cref{fig:pd_split}.
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/pd_split_thermal.pdf}
    \caption{\textbf{Qutrit phase space regions for $W_{\rho | \tau}(\bmz)$.}
    Here, the negative component of the magic state overlaps the Wigner distribution of $|0\>$. The rescaled distribution attains a single value in each of the four regions, proportional to the value depicted in the region, see~\cref{eq:bmw_rescaled}.
    }
    \label{fig:pd_split}
\end{figure}

@@

Applied to Wigner distributions $W_\rho (\z)$ for a quantum state $\rho$ we have
\begin{equation}
	H_\alpha(W_\rho) := \frac{1}{1-\alpha} \log \sum_{\bmz} W_{\rho}(\bmz)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $2a / (2b-1)$, for non-negative integers $a,b$. For the noisy Strange state $\rho = \rho_{\rm{S}}(\epsilon)$, this becomes
\begin{align}
	H_{\alpha}\left(W_{\rho_{\rm{S}}(\epsilon)}\right) = \frac{1}{1-\alpha}\log \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right] \nonumber\\
\end{align}
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.3]{figs/H_vs_eps_a.pdf}
    \caption{The R\'{e}nyi entropy $H_\alpha$ is increasing in $\alpha$ and $\epsilon$. First dashed contour from the right corresponds to $H_\alpha(\rho_{\rm{S}}(\epsilon)) = 0$. Specifically, $H_\alpha(\rho_{\rm{S}}(0)) = 0$ that implies $6^\alpha - 2^{\alpha} = 8$, with the approximate solution $\alpha \approx 1.31$.
    }
    \label{fig:H}
\end{figure}
We now show the following equivalence between Wigner negativity and negative R\'{e}nyi entropies.
\begingroup
\def\thetheorem{\ref{thm:HSchur}}
\begin{theorem}
	A quantum state $\rho$ has Wigner negativity if and only if $H_\alpha(W_\rho) < 0$ for some $\alpha =  \frac{2a}{2b-1}$, with positive integers $a \geq b$.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
\begin{proof} 
If we have $H_\alpha (W_\rho) <0 $ for $\alpha >1$ then this implies that $\log \sum_{\bmz} W_\rho (\bmz)^\alpha > 1$, however it is known that $H_\alpha$ is always non-negative on probability distributions and therefore $W_\rho(\bmz)$ must be a strict quasi-distribution with negativity.

>>>>

\begin{figure}[t]
    \centering
    \includegraphics[scale=0.35]{figs/distill_bounds.pdf}
    \caption{\textbf{Magic distillation bounds from majorization of Wigner quasi-distributions.} The above plot shows distillation upper bounds for qutrits obtained from majorisation on unital protocols. The constraints are plotted as a function of the depolarizing error $\epsilon$ for a noisy Strange state (Equation~\ref{eq:noisysn}). The bound $R_\infty$ follows from a Lorenz curve analysis presented in Section~\ref{sec:unital}, while the bound $R_{10}$ comes from single-shot entropies on the Wigner quasi-distributions, as discussed in Section~\ref{section:entropies}. Numerical bounds $R_{\rm{num}}$ are obtained by using the full set of majorization constraints. The new bounds $R_\infty$, $R_{10}$, $R_{\rm{num}}$ all provide stricter constraints than mana~\cite{cit:veitch2} and max--thauma $\theta_{\rm{max}}$~\cite{Wang_2020} bounds for noisy Strange states.
    }
    \label{fig:distill_bounds}
\end{figure}

@@

\subsection{Magic theories for quantum computation}
\label{sec:mono}
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3]{figs/protocol_diagram.pdf}
    \caption{\textbf{Probing magic protocols with reference states.} 
	A magic protocol (top figure) converts $n$ copies of some noisy magic state $\rho$ with noise parameter $\epsilon$ to $m$ copies of some less noisy state $\rho'$ with $\epsilon' < \epsilon$. The physics of the protocol can be analysed by considering how it would transform some distinguished reference stabilizer state $\tau^{\otimes n}$ (bottom figure). For the case where this reference state is a thermal state at some temperature $T$ our magic distillation bounds are functions of $T$, the error parameters $\epsilon, \epsilon'$, and \nick{the free energies $F, F'$ that the protocol channel $\E$ injects in}. In addition, a quantity $\phi$ (Equation~\ref{eq:phi}) links the computational and physical aspects, and corresponds to the degree to which the Wigner negativity of $\rho$ can be assigned a sharp energy via the system Hamiltonian.}
    \label{fig:sketch}
\end{figure}
A number of magic theories exist, where magic states are computational resource states with respect to a natural class of quantum operations that are considered free~\cite{Gour_2019}. One very natural class of free operations are those obtained from Clifford unitaries, Pauli measurements and the ability to discard quantum systems. However, there are also other candidates~\cite{cit:ahmadi, cit:seddon, Wang_2019}.

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.35]{figs/lc_strange.pdf}
    \caption{\textbf{A `heretical' family of Lorenz curves.} Traditionally, Lorenz curves for distributions are monotone increasing cumulant functions that reach a maximum value of $1$. In contrast, Lorenz curves for magic states break through the value of $L(x)=1$ due to the presence of negativity in the associated quasi-probability distributions. The above family of curves correspond to multiple copies of noisy Strange states $\rho\coloneqq\rho_{\rm{S}}(\epsilon)^{\otimes n}$ (Equation~\ref{eq:noisysn}) for $n=2,4,6$ within $\R_{\sigma}$, where $\sigma = \id/3$. Solid lines represent pure Strange states, while dashed lines represent $\epsilon$-noisy Strange states with depolarising error $\epsilon = 0.1$.
    }
    \label{fig:lcs}
\end{figure}

@@

\begin{figure}[t!]
    \centering
    \subfigure[][]{%
    \label{fig:noCliff}%
    \includegraphics[scale=0.35]{figs/rate_scatter.pdf}}\\
    \subfigure[][]{%
    \label{fig:Cliff}%
    \includegraphics[scale=0.35]{figs/rate_scatter_cliff.pdf}}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.} Shown are two contour plots of the bound on $R(\epsilon, \beta)$ for the case $H = H' = \sum_{k \in \mathbb{Z}_3} k \ketbra{k}$ and $\epsilon' = 0$, where $\beta$ is the inverse temperature and $\epsilon$ is the depolarising error of the input magic state. 
The top figure \subref{fig:noCliff} does not use any changes of Clifford basis, and the form of the bound depends on both the noise parameter and temperature. The curved dashed line is $\epsilon_\star(\beta)$ and given by Equation~\ref{eq:threshold} and $\beta_\star = (kT_\star)^{-1}$ is given by $E_2-\phi = kT_\star \ln 2$. 
In the bottom figure \subref{fig:Cliff} Clifford processing is used resulting in a smoother bound. In both figures the $\beta = 0$ line correspond to the unital bounds.}
    \label{fig:rate_contour}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.35]{figs/R_vs_A.pdf}
    \caption{\textbf{Variation of distillation bounds with Hamiltonian.} \nick{ We demonstrate non-trivial variation of our bound in Equation~\ref{eq:rate_bounds_proof} with repect to the system Hamiltonian. To this end, we fix $(\epsilon, \epsilon', \beta) = (0.1, 0.0, 0.2)$ and look at a family of Hamiltonians for the qutrit systems, given by $H = A_{\bmo}$ and $H' = (1-p-q)A_{\bmo}+pA_{(1,2)}+q{\rm{diag}}(0,1,2)$ for varying probabilities $p, q$. NEED TO DISCUSS VALUE $1.01$}
    }
    \label{fig:rvsa}
\end{figure}

@@

We can now apply this to obtain a family of magic distillation bounds based on R\'{e}nyi divergences.
\begin{theorem}\label{thm:entropic_bounds}
	Consider a general magic state distillation process on prime dimension qudits, that converts a magic state $\rho^{\otimes n} \longrightarrow \E(\rho^{\otimes n})=\rho'^{\otimes m}$and let $\tau$ be any full-rank stabilizer reference state on a qudit. Then the distillation rate $R := m/n$ is upper bounded as,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)}{D^{\rm reg}_\alpha( \rho', \tau')},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and with the regularized divergence
	\begin{equation}
D^{\rm reg}_\alpha( \rho', \tau') := \frac{1}{m} D_\alpha (W_{\rho'^{\otimes m}} || W_{\tau'_m}),
\end{equation}
between the output magic state $\rho'^{\otimes m}$ and $\tau'_m = \E(\tau^{\otimes n})$.
\end{theorem}
\nick{Notation $D^{\rm reg}_\alpha( \rho', \tau')$ is a little clunky. What are $\rho'$ and $\tau'$ ? Maybe at least use $D^{\rm reg}_\alpha( \rho', \tau'_m)$ as in the presence of correlations marginal state $\tau'$ is not well defined}
The proof is given in Appendix~\ref{app:single-shot}.

@@

If correlations in $\tau'_m$ between subsystems can be neglected then \nick{we can write $\tau'_m = \tau'^{\otimes m}$ for some sharp marginal state $\tau'$} and the regularised divergence reduces to
\begin{equation}
D^{\rm reg}_\alpha( \rho', \tau') = D_\alpha (W_{\rho'} || W_{\tau'}),
\end{equation}
which gives a generalized version of Theorem \ref{thm:free-energy}.

@@

The Wigner distribution of state $\tau$ is then given by
\begin{align}
	\W[\bmz]{\tau} &= \sum\limits_{k=0}^2 \frac{e^{-\beta E_k}}{\Z}W_{|k\>\<k|}(q, p) \nonumber\\
	&= \sum\limits_{k=0}^2 \frac{e^{-\beta E_k}}{\Z} \delta_{q, k} = \frac{e^{-\beta E_x}}{3\Z},
\end{align}
where $x$ labels one of the three vertical lines in the phase space.
The rescaled Wigner distribution $W_{\rho|\tau}(\bmz)$ can now be computed. It has $9$ components, but these come with multiplicities. In total, there are four distinct values on the phase space, as illustrated in~\cref{fig:pd_split}.
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/pd_split_thermal.pdf}
    \caption{\textbf{Qutrit phase space regions for $W_{\rho | \tau}(\bmz)$.}
    Here, the negative component of the magic state overlaps the Wigner distribution of $|0\>$. The rescaled distribution attains a single value in each of the four regions, proportional to the value depicted in the region, see~\cref{eq:bmw_rescaled}.
    }
    \label{fig:pd_split}
\end{figure}

@@

Applied to Wigner distributions $W_\rho (\z)$ for a quantum state $\rho$ we have
\begin{equation}
	H_\alpha(W_\rho) := \frac{1}{1-\alpha} \log \sum_{\bmz} W_{\rho}(\bmz)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $2a / (2b-1)$, for non-negative integers $a,b$. For the noisy Strange state $\rho = \rho_{\rm{S}}(\epsilon)$, this becomes
\begin{align}
	H_{\alpha}\left(W_{\rho_{\rm{S}}(\epsilon)}\right) = \frac{1}{1-\alpha}\log \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right] \nonumber\\
\end{align}
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.35]{figs/H_vs_eps_a.pdf}
    \caption{\textbf{R\'{e}nyi entropy $H_\alpha$ of $\epsilon$-noisy Strange state.} \nick{$H_\alpha$ is increasing in $\alpha$ and $\epsilon$. The first dashed contour from the right corresponds to $H_\alpha(\rho_{\rm{S}}(\epsilon)) = 0$. In particular, $H_\alpha(\rho_{\rm{S}}(0)) = 0$ corresponds to $\alpha \approx 1.31$.}
    }
    \label{fig:H}
\end{figure}
We now show the following equivalence between Wigner negativity and negative R\'{e}nyi entropies.
\begingroup
\def\thetheorem{\ref{thm:HSchur}}
\begin{theorem}
	A quantum state $\rho$ has Wigner negativity if and only if $H_\alpha(W_\rho) < 0$ for some $\alpha =  \frac{2a}{2b-1}$, with positive integers $a \geq b$.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
\begin{proof} 
If we have $H_\alpha (W_\rho) <0 $ for $\alpha >1$ then this implies that $\log \sum_{\bmz} W_\rho (\bmz)^\alpha > 1$, however it is known that $H_\alpha$ is always non-negative on probability distributions and therefore $W_\rho(\bmz)$ must be a strict quasi-distribution with negativity.
